Revisiting the Truisms of Software Engineering:\\ Does Phase Delay Dramatically Increase Repair Time?

### 2015-03-13 data.tex
dropped correaltion scatter plots. replaced pngs with pdfs

\subsection{Projects in the Sample}
Since 2004, one of the authors (Nichols) has been been mentoring 171 software development teams around the world as they deploy TSP within their organizations.  
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).
The majority of
the projects were either web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany),  a few from a commercial 
computer-aided design system (developed by a world wide distributed team), and less that  
a dozen United States
Department of Defence projects.
 An anonymized version of that data is available in the PROMISE repository\footnote{\carter{Need to add}} and,
 for confidentiality restrictions, we cannot offer 
further details
on those project.

@@

\subsection{Data Collection Process}
\sei{
\bi
    \item What are artifacts uploaded to the SEI tool?
    \item What are the units contained in those artifacts? Are they the ``issues'' as defined in the following paragraph?
    \item Is participation in the SEI data collection mandatory? i.e., is the data provided self selected?
    \item When is the data collected?
    \item What are screening/cleaning procedures, if any?
\ei
}

\sei{ wrn: please review the following}

Organizations using TSP agree to provide project data to the SEI for use in research with the provision that the data must not be traceable to its source. The data includes project and site  characteristic summaries, surveys of team members, launch presentations, launch outbriefs, and baseline plans, final data from the project, and a post mortem report.  In practice, this data requirement has only been enforced for the purposes of certifying and reauthorizing TSP coaches who are the only individuals authorized to submit data. To this data, only the data recorded using the Process Dashboard tool has been collected and aggregated for research.

@@

\subsection{Project Phases}
This paper studies the impact of phase delay on the time required to resolve issues.
The logical phases used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases have the similar  sub-phases of {\em review} and {\em inspect}. In TSP, individuals perform personal reviews on products prior to the peer inspection.\footnote{\bill{plz distinguish review and inspect. does one usually happen first?}}
\item Testing is divided into several statges \footnote{\bill{
(this appears to be OBE), need a one
line description of test vs qualtest vs inttest vs systemTest vs AcceptTest. I tries some words in the Key to \fig{sedata}. dont know if i got it right.}}
\ei

\todo{I recommend deleting the rest of this section. Replace with better descriptions of what the phases are.}
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

>>>>

\subsection{Projects in the Sample}
Since 2004, one of the authors (Nichols) has been been mentoring software development teams and coaches around the world as they deploy TSP within their organizations.  
The projects
were mostly small to medium, with a median duration of 46 days
and a maximum duration of 90 days. Median team size was 7
people, with a maximum of 40. Hence,  total development 
effort\footnote{
Effort =  duration*teamSize/workDaysPerYear.} for
these projects ranged from 1.3 years (median) to 14.3 years (max).
The majority of
the projects were either web portals or banking systems in the US, South Africa, and Mexico. 
There were also some  medical devices (US, France, Japan, and Germany),  a few from a commercial 
computer-aided design system (developed by a world wide distributed team.
 An anonymized version of that data is available in the PROMISE repository\footnote{\carter{Need to add}} and,
 for confidentiality restrictions, we cannot offer 
further details
on those project.

@@

\subsection{Data Collection Process}

Organizations using a TSP agree to provide their project data to the SEI for use in research. In return the SEI agrees    that  data must not be traceable to its source. The data is collected at major project events, launch, interim checkpoints, and at completion. Data includes project and site  characteristic summaries, surveys of team members, launch presentations, launch outbriefs, and baseline plans, final data from the project, and the post mortem report.  In practice, this data requirement has only been enforced for the purposes of certifying and reauthorizing TSP coaches who are the only individuals authorized to submit data. To this data, only the data recorded using the Process Dashboard tool has been collected and aggregated for research.

@@

\subsection{Project Phases}
This paper studies the impact of phase delay on the time required to resolve issues.
The logical phases used in this paper are shown in \fig{waterfall}. Note that, in that figure:
\bi 
\item
Several  phases in which product is created have sub-phases of {\em review} and {\em inspect} to remove defects. TSP uses review when individuals perform personal reviews of their work products prior to the peer review which TSP calls the inspection).
\item Testing is divided into several stages. Developers perform unit test prior to code complete. Some standard test phases after code complete include the integration which combines program units into workable system ready for system test. Integration and system test are often performed by a separate group.

\todo{I recommend deleting the rest of this section. Replace with better descriptions of what the phases are.
wrn: I agree, I will do this Friday}
Also shown in \fig{waterfall} (bottom left) is the standard definition of an agile process~\cite{boehmturner03}:
Given some
some backlog of tasks, wgeb teams complete their current tasks, they select the next task(s) to complete. That selection process may use a variety of criteria
to prioritize which  tasks are selected (for more details on that selection process, see~\cite{me09j,port08,boehmturner03}). Tasks are completed in ``sprints'' that can last hours,
days, but rarely not more than weeks. Each day meet for brief ``scrum'' sessions to assess (and possibly alter) their current progress on the goals of the sprint.  
Agile teams race to generate releases
(in the continuous release model, releases can be generated on a daily basis, or even faster).  
Experience gained from those releases informs the discussion in the daily scrums which, in turn,
can inform the team's decisions on how to select and implement the next set of tasks for next sprint.
In fact, that experience can result in changing some/all of the tasks in the current backlog.

### 2015-03-13 phaseDelay.tex
dropped correaltion scatter plots. replaced pngs with pdfs

\todo{
\bi
    \item What is the cost-to-fix curve in TSP data? Check against H1 and compare to past literature.
    \item What is the effect of phase delay? Check against H2.
\ei
}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3.3in]{fix-phase-dist.png}
\end{center}
\caption{Distribution of defects found and fixed by phase.}
\label{fig:fix-phase-dist}
\end{figure}

@@

\subsection{Phase delay analysis}
\todo{
\bi
    \item What can we say about \fig{raw}
    \item What can we say about \fig{scale}
    \item What is the difference between \fig{raw} and \fig{scale}
\ei
}

 \fig{phase_fixtime_plot} shows scatterplots of the cost-to-fix (rank) per phase against the phase delay. In the early phases, i.e., Design \& Coding, most of the defects found have a relatively small phase delay. This is ideal considering that the purpose of the early phase reviews is to remove as many defects as possible.

In the TSP data, there is a weak correlation (Kendall's $\tau = 0.19, p < 0.01, 95\% CI$) between a defect's cost-to-fix (minutes of effort) and the defects phase delay (phase removed - phase injected) - see \fig{correlation}. The correlations were also calculated for each phase, e.g., does the phase delay of defects found in Test correlate with the defect's cost-to-fix (\fig{phase_corr}).

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3.3in]{corr.png}
\end{center}
\caption{Correlation of phase delay and fix time (plot shows a 1\% sampling of data).}
\label{fig:correlation}
\end{figure}

\begin{figure}[ht]
\centering
\begin{tabular}{lll}
  Phase & $tau$ & $p$ \\ 
  \hline
  Design & 0.14 & 0.08 \\ 
  DesignReview & 0.06 & $<$0.01 \\ 
  DesignInspect & 0.05 & $<$0.01 \\ 
  Code & 0.27 & $<$0.01 \\ 
  CodeReview & 0.12 & $<$0.01 \\ 
  CodeInspect & 0.09 & $<$0.01 \\ 
  Test & 0.13 & $<$0.01 \\ 
  IntTest & 0.1 & $<$0.01 \\ 
  SysTest & 0.25 & $<$0.01 \\ 
  AcceptTest & 0.06 & 0.29 \\ 
  \end{tabular}
\caption{Correlation of cost-to-fix and phase delay for defects found (Kendall's $\tau$)} 
\label{fig:phase_corr}
\end{figure}

\begin{figure*}[!ht]
\centering
\includegraphics[width=1.6in]{corr-Code.png}
\includegraphics[width=1.6in]{corr-CodeReview.png}
\includegraphics[width=1.6in]{corr-CodeInspect.png}
\vspace{0.01cm}

\includegraphics[width=1.6in]{corr-Design.png}
\includegraphics[width=1.6in]{corr-DesignReview.png}
\includegraphics[width=1.6in]{corr-DesignInspect.png}
\vspace{0.01cm}

\includegraphics[width=1.6in]{corr-Test.png}
\includegraphics[width=1.6in]{corr-IntTest.png}
\includegraphics[width=1.6in]{corr-SysTest.png}
\includegraphics[width=1.6in]{corr-AcceptTest.png}

\caption{Correlation of cost-to-fix (rank) and phase delay for defects found by phase for phases where \# defects >= 30 and \# unique phase delay values >= 5. x = phase delay, y = cost-to-fix rank. Plots use random samples from each phase for readability.}
\label{fig:phase_fixtime_plot}
\end{figure*}

@@

 \begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{median}\\
  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\input{deltas}
\end{tabular}
\end{center}
\caption{50th percentile (median) scale ups  for  time to resolve issues (taken from \fig{raw}).}
\label{fig:scale}
\end{figure}

\subsection{Phase Delay in Slower Bugs?}
In these results, we checked if a small number of bugs are most expensive. 
To check this:
\bi
\item
We repeated the analysis that generated \fig{scale},
but instead of looking at the 50th percentile, we displayed the scale up factors
\item 
We only
checked the Design and Code scale up results since, from \fig{scale}, it is clear these
have the most examples of longest phase delays.
\ei 
The  results are shown in \fig{scale90} and these
results are somewhat different for Design and Coding issues: 
\bi 
\item For Design issues,  these have the same
general form as the 50th percentile results. That is, while it it certainly faster
to remove thing sin the phase where they are created, once we leave that phase
it does not seem to matter much how many phases we wait before fixing the issue.
\item
\item For Coding issues, the seems little impact of phase delay on the time
required to fix.
\ei

\begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
          \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{99th }\\
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{percentile }\\
  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\input{deltasBigBig}
\end{tabular}
\end{center}
\caption{99th percentile scale ups.}
\label{fig:scale99}
\end{figure}

\begin{figure}[!t]
\renewcommand{\baselinestretch}{0.7}
\scriptsize
\begin{center}
\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{mean}\\
  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
\input{deltasMean}
\end{tabular}
\end{center}
\caption{Mean  scale ups.}
\label{fig:scale90}
\end{figure}

@@

\section*{Acknowledgements}
This work was partially funded by an National Science
Foundation grants NSF-CISE 1302169 and CISE 1302169.

>>>>

\todo{
\bi
    \item What is the cost-to-fix curve in TSP data? Check against H1 and compare to past literature.
    \item What is the effect of phase delay? Check against H2.
    \item fig12 is great. really important. its from  different decades, right?  so can we add text saying that the general pattern is that the (approx) 15 years between each curve tends to a flattening. not, perhaps, due to process but to better build environs.
    \item but fig12 is blurred in the words. recreate using crisper fonts?
    \item fig 10 text a little blurred. can we re snap that one? or generate it it some higher res form?
    \item figs 13 (fix time rank v phase delay) is hard to read (text)
\ei
}

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3.3in]{fix-phase-dist.pdf}
\end{center}
\caption{Distribution of defects found and fixed by phase.}
\label{fig:fix-phase-dist}
\end{figure}

@@

\subsection{Phase delay analysis}
\todo{
\bi
    \item What can we say about \fig{raw}
    \item What is the difference between \fig{raw}
\ei
}

 In the early phases, i.e., Design \& Coding, most of the defects found have a relatively small phase delay. This is ideal considering that the purpose of the early phase reviews is to remove as many defects as possible.

In the TSP data, there is a weak correlation (Kendall's $\tau = 0.19, p < 0.01, 95\% CI$) between a defect's cost-to-fix (minutes of effort) and the defects phase delay (phase removed - phase injected) - see \fig{correlation}.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3.3in]{corr.pdf}
\end{center}
\caption{Correlation of phase delay and fix time (plot shows a 1\% sampling of data).}
\label{fig:correlation}
\end{figure}

% \begin{figure}[ht]
% \centering
% \begin{tabular}{lll}
%   Phase & $tau$ & $p$ \\ 
%   \hline
%   Design & 0.14 & 0.08 \\ 
%   DesignReview & 0.06 & $<$0.01 \\ 
%   DesignInspect & 0.05 & $<$0.01 \\ 
%   Code & 0.27 & $<$0.01 \\ 
%   CodeReview & 0.12 & $<$0.01 \\ 
%   CodeInspect & 0.09 & $<$0.01 \\ 
%   Test & 0.13 & $<$0.01 \\ 
%   IntTest & 0.1 & $<$0.01 \\ 
%   SysTest & 0.25 & $<$0.01 \\ 
%   AcceptTest & 0.06 & 0.29 \\ 
%   \end{tabular}
% \caption{Correlation of cost-to-fix and phase delay for defects found (Kendall's $\tau$)} 
% \label{fig:phase_corr}
% \end{figure}

% \begin{figure*}[!ht]
% \centering
% \includegraphics[width=1.6in]{corr-Code.png}
% \includegraphics[width=1.6in]{corr-CodeReview.png}
% \includegraphics[width=1.6in]{corr-CodeInspect.png}
% \vspace{0.01cm}

% \includegraphics[width=1.6in]{corr-Design.png}
% \includegraphics[width=1.6in]{corr-DesignReview.png}
% \includegraphics[width=1.6in]{corr-DesignInspect.png}
% \vspace{0.01cm}

% \includegraphics[width=1.6in]{corr-Test.png}
% \includegraphics[width=1.6in]{corr-IntTest.png}
% \includegraphics[width=1.6in]{corr-SysTest.png}
% \includegraphics[width=1.6in]{corr-AcceptTest.png}

% \caption{Correlation of cost-to-fix (rank) and phase delay for defects found by phase for phases where \# defects >= 30 and \# unique phase delay values >= 5. x = phase delay, y = cost-to-fix rank. Plots use random samples from each phase for readability.}
% \label{fig:phase_fixtime_plot}
% \end{figure*}

@@

 %\begin{figure}[!t]
%\%renewcommand{\baselinestretch}{0.7}
%\scriptsize
%\begin{center}
%\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
%%           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{median}\\
 % Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
%\input{deltas}
%\end{tabular}
%%\end{center}
%\caption{50th percentile (median) scale ups  for  time to resolve issues (taken from \fig{raw}).}
%%\label{fig:scale}
%\end{figure}

%\subsection{Phase Delay in Slower Bugs?}
%In these results, we checked if a small number of bugs are most expensive. 
%To check this:
%\bi
%\item
%We repeated the analysis that generated \fig{scale},
%but instead of looking at the 50th percentile, we displayed the scale up factors
%\item 
%We only
%checked the Design and Code scale up results since, from \fig{scale}, it is clear these
%have the most examples of longest phase delays.
%\ei 
%The  results are shown in \fig{scale90} and these
%results are somewhat different for Design and Coding issues: 
%\bi 
%\item For Design issues,  these have the same
%general form as the 50th percentile results. That is, while it it certainly faster
%to remove thing sin the phase where they are created, once we leave that phase
%it does not seem to matter much how many phases we wait before fixing the issue.
%\item
%\item For Coding issues, the seems little impact of phase delay on the time
%required to fix.
%\ei

%\begin{figure}[!t]
%\renewcommand{\baselinestretch}{0.7}
%\scriptsize
%\begin{center}
%\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
%          \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{99th }\\
%           \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{percentile }\\
%  Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
%\input{deltasBigBig}
%\end{tabular}
%\end{center}
%%\caption{99th percentile scale ups.}
%\label{fig:scale99}
%\end{figure}

%\begin{figure}[!t]
%\renewcommand{\baselinestretch}{0.7}
%\scriptsize
%\begin{center}
%\begin{tabular}{l@{~~}|l@{~}|r@{~}|r@{~}r@{~}|r@{~}l}
 %          \multicolumn{2}{c}{~}                 &  &\multicolumn{2}{c|}{mean}\\
 % Injection&   Removal& $n$ & initial & now & \multicolumn{2}{l}{Scale up, w.r.t. initial}
%\input{deltasMean}
%\end{tabular}
%\end{center}
%\caption{Mean  scale ups.}
%\label{fig:scale90}
%\end{figure}

@@

\section*{Acknowledgements}
This work was partially funded by an National Science
Foundation grants NSF-CISE 1302169 and CISE 1302169. 
\\
We thank Tuma Solutions for providing the Team Process Data Warehouse software in support of this research.
